---
title: Run calculation and extraction functions
author: "Miguel Rodo"
format:
  html:
    embed-resources: true
---

```{r}
#| include: false
library(tibble)
library(ggplot2)
library(dplyr)
for (x in list.files("R", full.names = TRUE)) {
  source(x)
}
```


## Calculation

Here we load the processed data:

```{r}
# this function is defined in R/load.R
# (and sourced in the initial chunk above)
path_data_tidy_vacc_freq <- path_get_data_tidy_vacc_freq()
load(path_data_tidy_vacc_freq)
```

```{r}
data_tidy_vacc_freq
```

### Vaccine-induced response

To calculate the vaccine-induced response, we subtract the baseline response.
If the baseline response is missing for a particular individual, we subtract the 
median baseline response for that individual's vaccine and infection status group.

```{r}
data_tidy_vacc_raw <- data_tidy_vacc_freq
data_tidy_vacc_vi <- data_tidy_vacc_freq |>
  vacc_calc_vaccine_induced_response(cn_resp = "response")
data_tidy_vacc_vi
```

### Summed response

To calculate the summed response, we sum over the responses
for a given individual at a given timepoint.
The `vacc_calc_response_summed` function performs this:

```{r}
data_tidy_vacc_raw_summed <- data_tidy_vacc_raw |>
  vacc_calc_response_summed(cn_resp = "response")
data_tidy_vacc_vi_summed <- data_tidy_vacc_vi |>
  vacc_calc_response_summed(cn_resp = "response")
data_tidy_vacc_vi_summed
```

### Profile response

To calculate the profile response, we divide the response
(ito a given cytokine combination) for a given
individual at a given timepoint by the sum of the
absolute values of their responses at that timepoint.
The `vacc_calc_response_profile` function performs this:

```{r}
data_tidy_vacc_raw_profile <- data_tidy_vacc_freq |>
  vacc_calc_response_profile(cn_resp = "response")
data_tidy_vacc_vi_profile <- data_tidy_vacc_vi |>
  vacc_calc_response_profile(cn_resp = "response")
data_tidy_vacc_vi_profile
```

## Subset

The functions `vacc_extract_baseline`, `vacc_extract_peak`, and `vacc_extract_memory`
extract the baseline, peak, and memory responses, respectively.

The baseline response is the response at day 0.
The peak response is the maximum response across all timepoints (consistent across individuals within a vaccine).
The memory response is the response at the last timepoint.

### Baseline

```{r}
data_tidy_vacc_raw_summed_baseline <- data_tidy_vacc_raw_summed |>
  vacc_extract_baseline()
data_tidy_vacc_raw_summed_baseline
data_tidy_vacc_raw_profile_baseline <- data_tidy_vacc_raw_profile |>
  vacc_extract_baseline()
data_tidy_vacc_raw_profile
```

### Peak

```{r}
data_tidy_vacc_raw_summed_peak <- data_tidy_vacc_raw_summed |>
  vacc_extract_peak()
data_tidy_vacc_raw_summed_peak
data_tidy_vacc_raw_profile_peak <- data_tidy_vacc_raw_profile |>
  vacc_extract_peak()
data_tidy_vacc_raw_profile_peak
```

#### Memory

```{r}
data_tidy_vacc_raw_summed_mem <- data_tidy_vacc_raw_summed |>
  vacc_extract_memory()
data_tidy_vacc_raw_summed_mem
data_tidy_vacc_raw_profile_mem <- data_tidy_vacc_raw_profile |>
  vacc_extract_memory()
data_tidy_vacc_raw_profile_mem
```
