---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tibble)
```

# DataTidy24TBVaccSexDiff

<!-- badges: start -->
<!-- badges: end -->

The purpose of `DataTidy24TBVaccSexDiff` is to generate analysis-ready data for the assessment of sex-associated differences in TB vaccine-induced antigen-specific CD4 T cell responses.

## Installation

To install `DataTidy24TBVaccSexDiff`, run:

```{r echo = TRUE, eval = FALSE}
remotes::install_github(
  "SATVILab/DataTidy24TBVaccSexDiff"
)
```

## Usage

`DataTidy24TBVaccSexDiff` provides processed data, as well as convenience functions for calculating transformations of the response (summed and profile; see Rodo (2019) for details) and extracting relevant timepoints (baseline, peak and memory; again, see Rodo (2019) for details).

### Loading data

To load the data, run:

```{r}
library(DataTidy24TBVaccSexDiff)
data("data_tidy_vacc_freq")
data_tidy_vacc_freq
```

## Calculating summed and profile responses

To calculate the summed response, run:

```{r}
vacc_calc_response_summed(
  data_tidy_vacc_freq, "response"
)
```

To calculate the summed profile, run:

```{r}
vacc_calc_response_profile(
  data_tidy_vacc_freq, "response"
)
```

## Extracting timepoints

To extract the baseline response, run:

```{r}
vacc_extract_baseline(data_tidy_vacc_freq)
```

To extract the peak response, run:

```{r}
vacc_extract_peak(data_tidy_vacc_freq)
```

To extract the memory response, run:

```{r}
vacc_extract_memory(data_tidy_vacc_freq)
```

Note that these work automatically after the calculation of summed or profile responses (and the order is irrelevant), for example:

```{r}
data_tidy_vacc_freq |>
  vacc_calc_response_summed("response") |>
  vacc_extract_peak()
data_tidy_vacc_freq |>
  vacc_extract_baseline() |>
  vacc_calc_response_profile("response")
```

In addition, a convenience function for setting negative responses to zero is provided:

```{r}
vacc_set_neg_to_zero(data_tidy_vacc_freq, "response")
```

## Reproduction

*To be completed*
